---
title: "Activity assignment"
author: "Hyun Seok Kwak"
date: "2025-03-02"
output: html_document
---

## Markdown setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_TIME", "en_US.UTF-8") # set local time
```

## Library

```{r library}
lib <- c("dplyr", "ggplot2")
for(i in lib){
if (!requireNamespace(i, quietly = TRUE)) {
  install.packages(i)
}}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data

```{r data loading}
if (!file.exists("activity.csv")) {
  fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file(fileUrl, destfile = "repdata_data_activity.zip")
  unzip("repdata_data_activity.zip")
}

data <- read.csv("./activity.csv",
                 header = TRUE)
```

Structure and summary of dataset:
```{r}
print(str(data))
print(summary(data))
```


## What is mean total number of steps taken per day?

Aggregate steps by date:
```{r}
total_steps <- with(data, 
                    aggregate(steps, by = list(date), sum, na.rm = TRUE))
names(total_steps) <- c("date", "steps")
```


Histogram using ggplot:
```{r}
hist <- ggplot(aes(x = steps),
               data = total_steps) +
  geom_histogram(breaks = seq(0, 25000, by = 2500),
                 fill = "lightblue", col = "black") +
  ggtitle("Total Number of Steps Taken on a Day") +
  xlab("Average steps per day") +
  ylab("Frequency") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
print(hist)
```

Median number of steps taken each day:
```{r}
median(total_steps$steps)
```

Mean number of steps taken each day:
```{r}
mean(total_steps$steps)
```


## What is the average daily activity pattern?

Aggregate steps by interval:
```{r}
interval_steps <- with(data, 
                       aggregate(steps, by = list(interval), 
                                 mean, na.rm = TRUE))
names(interval_steps) <- c("interval", "steps")
```


Line graph with ggplot:
```{r}
plot <- ggplot(aes(x = interval, y = steps),
               data = interval_steps) +
  geom_line(col = "blue") +
  ggtitle("Average Number of Steps per interval") +
  xlab("Interval") +
  ylab("Average steps") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
print(plot)
```


Interval with maximum number of average steps:
```{r}
interval_steps[which.max(interval_steps$steps),]$interval
```


## Imputing missing values

Total number of missing values in the dataset:
```{r}
sum(is.na(data$steps))
```

Impute missing value as mean of average steps per interval:
```{r}
data_impute <- data %>% 
  left_join(interval_steps, by = "interval") %>% 
  mutate(steps = ifelse(is.na(steps.x), steps.y, steps.x)) %>% 
  select(-steps.x, -steps.y)
```


Aggregate steps by date:
```{r}
total_steps2 <- with(data_impute, 
                     aggregate(steps, by = list(date), sum, na.rm = TRUE))
names(total_steps2) <- c("date", "steps")
```

Plot a histogram with ggplot:
```{r}
hist2 <- ggplot(aes(x = steps),
               data = total_steps2) +
  geom_histogram(breaks = seq(0, 25000, by = 2500),
                 fill = "lightblue", col = "black") +
  ggtitle("Total Number of Steps Taken on a Day") +
  xlab("Average steps per day") +
  ylab("Frequency") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
print(hist2)
```

Median number of steps taken each day:
```{r}
median(total_steps2$steps)
```

Mean number of steps taken each day:
```{r}
mean(total_steps2$steps)
```

Compared to the estimates from the first part of the assignment, imputing missing value increases mean and median of total number of steps per day


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variables indicating weekday or weekend:
```{r}
data_impute <- data_impute %>% 
  mutate(type = ifelse(weekdays(as.Date(date)) %in% c("Saturday", "Sunday"),
                       "weekend", "weekdays")) %>% 
  mutate(type = factor(type, levels = c("weekdays", "weekend")))
```


Aggregate steps by interval:
```{r}
interval_steps2 <- with(data_impute, 
                       aggregate(steps, by = list(interval, type), 
                                 mean, na.rm = TRUE))
names(interval_steps2) <- c("interval", "type", "steps")
```


Pannel plot with ggplot:
```{r}
plot2 <- ggplot(aes(x = interval, y = steps, col = type),
               data = interval_steps2) +
  geom_line() +
  facet_wrap(~ type, ncol = 1, nrow = 2) +
  theme_classic()
print(plot2)
```